A GPU Sonar Simulator for Automatic Target Recognition

### 2014-05-23 Simulator/article.tex
la2pdf commit - 2014:05:23	16:36:55	(Fri23.May)
la2pdf commit - 2014:05:23	18:00:53	(Fri23.May)

We have developed a synthetic aperture sonar (SAS) simulator that avoids this problem by running sufficiently fast to create templates adapted to the actual scene as part of the classification process. This is achieved with the aid of the massive computing power in graphics processing units (GPUs) and optimized software libraries for scene rendering. The simulator loads a 3D model of the seafloor and an object class model into OpenGL, where emitted sound waves are modeled with a light source placed at the sonar transmit location. We assume rough, isotropic surfaces reflecting sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Blake1993,Bell1995}, where the intensity of the backscattered sound only depends on the incidence angle of the transmitted signal onto the model surface. When rendering OpenGL is set up to produce an optical 2D intensity image and depth map that reveals the distance from each pixel to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.

@@

The next step is to scale, orient and position the objects in the scene, and finally project this scene onto an image plane. This is achieved by applying a set of transformations to each vertex in the model. In OpenGL, the set of transformations are usually formalized like this:
\begin{align*}
\bmat{x\\y\\0\\1}_\text{image} = \P \cdot \V \cdot \M \cdot \bmat{x\\y\\z\\1}_\text{model},
\end{align*}
where $\bmat{\cdot}_\text{model}$ is the loaded model, $\bmat{\cdot}_\text{image}$ is the output image, and $\M$, $\V$ and $\P$ are transformation matrices. These will be described next.

The first transform we apply is the model matrix $\M$ which operates in the model's local coordinate system. It is defined as
\begin{align}
\bmat{x_t \\ y_t \\ z_t \\ w}_\text{\parbox{1cm}{\setlength\baselineskip{0.3cm}transformed\\model}} &= \boldsymbol{T} \cdot \R \cdot \boldsymbol{S} \bmat{x \\ y \\ z \\ w}_\text{model}.
\end{align}
where $\boldsymbol{S}, \R$ and $\boldsymbol{T}$ are matrices for scaling, rotation and translation, respectively. These are described in Appendix \ref{transformation_matrices}.

Next we map the model into world coordinates using the view matrix $\V$. This involves positioning and orienting the model relative to the camera. In our case we let the camera be the phase center of the sonar, and define it as the world origin. The transformation take the form of
\begin{align}
\V &= \mat{T}(0, dy, dz) \cdot \R_\text{x}(\phi_x, 0, 0) \nonumber\\
&= 
\bmat{
1  &  0  &  0  &  0 \\
0  &  1  &  0  &  dy \\
0  &  0  &  1  &  dz \\
0  &  0  &  0  &  1 \\
} \cdot
\bmat{
1  &  0           &  0           &  0 \\
0  &  \cos\phi_x  &  -\sin\phi_x &  0 \\
0  &  \sin\phi_x  &  \cos\phi_x  &  0 \\
0  &  0           &  0           &  1 \\
},
\end{align}
where
\begin{align*}
dy &= \\
dz &= \\
\phi_x &=
\end{align*}

@@

We implement rotations in a 3D euclidean space by combining matrices that rotate around each of the three respective dimensions.. For rotation by an angle $\phi_x$ clockwise around the $x$-axis we have:
\begin{align}
\R_x(\phi_x) &= \bmat{
1  &  0           &  0           &  0 \\
0  &  \cos\phi_x  &  -\sin\phi_x &  0 \\
0  &  \sin\phi_x  &  \cos\phi_x  &  0 \\
0  &  0           &  0           &  1 \\
},
\end{align}
and for the other two euclidean dimensions we have:
\begin{align}
\R_y(\phi_y) &= \bmat{
\cos\phi_y  &  0  &  \sin\phi_y  &  0 \\
0           &  1  &  0           &  0 \\
-\sin\phi_y &  0  &  \cos\phi_y  &  0 \\
0           &  0  &  0           &  1 \\
},
\end{align}
and
\begin{align}
\R_z(\phi_z) &= \bmat{
\cos\phi_z  &  -\sin\phi_z &  0  &  0 \\
\sin\phi_z  &  \cos\phi_z  &  0  &  0 \\
0           &  0           &  1  &  0 \\
0           &  0           &  0  &  1 \\
}.
\end{align}
The order in which these are applied matters. We rotate around the $x$-axis first, then around $y$ and finally around $z$:
\begin{align}
\R(\phi_x, \phi_y, \phi_z) &= \R_z(\phi_z)\;\R_y(\phi_y)\;\R_x(\phi_x).
\end{align}

>>>>

We have developed a synthetic aperture sonar (SAS) simulator that avoids this problem by running sufficiently fast to create templates adapted to the actual scene as part of the classification process. This is achieved with the aid of the massive computing power in graphics processing units (GPUs) and optimized software libraries for scene rendering. The simulator loads a 3D model of the seafloor and an object class model into OpenGL, where emitted sound waves are modeled with a light source placed at the sonar transmit location. We assume rough, isotropic surfaces reflecting sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Blake1993,Bell1995}, where the intensity of the backscattered sound solely depends on the incidence angle of the transmitted signal onto the model surface. When rendering OpenGL is set up to produce an optical 2D intensity image and depth map that reveals the distance from each pixel to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.

@@

The next step is to scale, orient and position the objects in the scene, and finally project this scene onto an image plane. This is achieved by applying a set of transformations to each vertex in the model. In OpenGL, the set of transformations are usually formalized as
\begin{align*}
\bmat{x\\y\\0\\1}_\text{image} = \P \cdot \V \cdot \M \cdot \bmat{x\\y\\z\\1}_\text{model},
\end{align*}
where $\bmat{\cdot}_\text{model}$ is the loaded model, $\bmat{\cdot}_\text{image}$ is the output image, and $\M$, $\V$ and $\P$ are transformation matrices. These will be described next.

The first transform we apply is the model matrix $\M$, which operates in the model's local coordinate system. It is defined as
\begin{align}
\bmat{x_t \\ y_t \\ z_t \\ w}_\text{\parbox{1cm}{\setlength\baselineskip{0.3cm}transformed\\model}} &= \boldsymbol{T} \cdot \R \cdot \boldsymbol{S} \bmat{x \\ y \\ z \\ w}_\text{model},
\end{align}
where $\boldsymbol{S}, \R$ and $\boldsymbol{T}$ are matrices for scaling, rotation and translation, respectively. These are described in Appendix \ref{transformation_matrices}.

Next we map the model into world coordinates using the view matrix $\V$. This involves positioning and orienting the model relative to the camera. In our case we let the camera be the phase center of the sonar, and define it as the world origin. The transformation take the form of
\begin{align}
\V &= \mat{T}(0, dy, dz) \cdot \R_\text{x}(\phi_x, 0, 0) \nonumber\\
&= 
\bmat{
1  &  0  &  0  &  0 \\
0  &  1  &  0  &  dy \\
0  &  0  &  1  &  dz \\
0  &  0  &  0  &  1 \\
} \cdot
\bmat{
1  &  0           &  0           &  0 \\
0  &  \cos\phi_x  &  -\sin\phi_x &  0 \\
0  &  \sin\phi_x  &  \cos\phi_x  &  0 \\
0  &  0           &  0           &  1 \\
},
\end{align}
where
\begin{align*}
dy &= \\
dz &= \\
\phi_x &= \arctan\left(\frac{y_\text{mean}}{x_\text{mean}}\right)
\end{align*}

@@

We implement rotations in a 3D euclidean space by combining matrices that rotate around each of the three respective dimensions.. For rotation by an angle $\phi_x$ clockwise around the $x$-axis we have
\begin{align}
\R_x(\phi_x) &= \bmat{
1  &  0           &  0           &  0 \\
0  &  \cos\phi_x  &  -\sin\phi_x &  0 \\
0  &  \sin\phi_x  &  \cos\phi_x  &  0 \\
0  &  0           &  0           &  1 \\
},
\end{align}
and similarly, for the other two euclidean dimensions we have
\begin{align}
\R_y(\phi_y) &= \bmat{
\cos\phi_y  &  0  &  \sin\phi_y  &  0 \\
0           &  1  &  0           &  0 \\
-\sin\phi_y &  0  &  \cos\phi_y  &  0 \\
0           &  0  &  0           &  1 \\
},
\end{align}
and
\begin{align}
\R_z(\phi_z) &= \bmat{
\cos\phi_z  &  -\sin\phi_z &  0  &  0 \\
\sin\phi_z  &  \cos\phi_z  &  0  &  0 \\
0           &  0           &  1  &  0 \\
0           &  0           &  0  &  1 \\
}.
\end{align}
The order in which these are applied matters. We rotate around the $x$-axis first, then around $y$ and finally around $z$:
\begin{align}
\R(\phi_x, \phi_y, \phi_z) &= \R_z(\phi_z)\;\R_y(\phi_y)\;\R_x(\phi_x).
\end{align}
